Probabilistic Catalogs for Crowded Stellar Fields

### 2012-11-20 paper/ms.tex
Minor editing

\begin{table}\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter & Value (Test Case 1) & Value (Test Case 2)\\
\hline
$N$ & 100 & 1000\\
$h_1$ & 0.3 & 0.3\\
$h_2$ & 0.6 & 0.6\\
$\alpha_1$ & 1.1 & 1.1\\
$\alpha_2$ & 2 & 2\\
\hline
$\sigma_0$ & 0.05 & 0.3\\
$\eta$ & 0 & 0 \\
$s_1$ & 0.02 & 0.02\\
$s_2$ & 0.1 & 0.1\\
$w$ & 0.5 & 0.5\\
\hline
\end{tabular}
\end{center}
\caption{True parameter values used to generate the simulated data. The main
difference between the two test cases is the number of stars and the noise level.
\label{tab:truth}}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/test_cases.eps}
\caption{Two simulated images.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\subsection{Test Case 2}
Test Case 2 is more challenging than Test Case 1 because the image contains
more stars. This increases the size of the computational task in a number of
ways: firstly, there will be more unknown parameters to infer, so the
convergence of the MCMC algorithm will take longer. Secondly, the time taken
to compute the predicted image from a catalog (in order to evaluate the
likelihood) is longer because of the larger number of stars. Using DNS, some
samples from the posterior distribution can be obtained in about a day on a
modern PC. However, this problem does not have a first-order phase transition
like Test Case 1. Therefore, it is tractable to run standard Metropolis-Hastings
MCMC targeting the posterior distribution.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/summaries.eps}
\end{center}
\caption{Summary images for test case 2. The left panel shows the posterior
mean high-resolution scene. The middle panel shows the posterior mean scene
when observed at the resolution of the data, and the right hand panel shows the
model residuals.
\label{fig:summaries}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_N.eps}
\end{center}
\caption{The number of stars in the catalogs produced by \sex~compared with the
actual number of stars in the images. \sex~finds only a small fraction of the
stars actually present. This is to be expected in the high-$\sigma$ case as
\sex~then only searches for stars above a certain flux, but at the low-$\sigma$
end the number of stars is still drastically underestimated. Note that the true
numbers
of stars shown in these plots are$\sim$82\% of the values listed in
Figure~\ref{tab:truth} because here we only count those stars with positions
in $[-1, 1] \times [-1, 1]$.
\label{fig:sex_N}}
\end{figure}

>>>>

\begin{table}\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter & Value (Test Case 1) & Value (Test Case 2)\\
\hline
$N$ & 100 & 1000\\
$h_1$ & 0.3 & 0.3\\
$h_2$ & 0.6 & 0.6\\
$\alpha_1$ & 1.1 & 1.1\\
$\alpha_2$ & 2 & 2\\
\hline
$\sigma_0$ & 0.05 & 0.3\\
$\eta$ & 0 & 0 \\
$s_1$ & 0.02 & 0.02\\
$s_2$ & 0.1 & 0.1\\
$w$ & 0.5 & 0.5\\
\hline
\end{tabular}
\end{center}
\caption{True parameter values used to generate the simulated data. The main
difference between the two test cases is the number of stars and the noise
level, both of which are higher in test case 2.
\label{tab:truth}}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/test_cases.eps}
\caption{The two simulated images used to test our methodology.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\subsection{Test Case 2}
Test Case 2 is more challenging than Test Case 1 because the image contains
more stars. This increases the size of the computational task in a number of
ways: firstly, there will be more unknown parameters to infer, so the
convergence of the MCMC algorithm will take longer. Secondly, the time taken
to compute the predicted image from a catalog (in order to evaluate the
likelihood) is longer because of the larger number of stars. Using DNS, some
samples from the posterior distribution can be obtained in about a day on a
modern PC. However, this problem does not have a first-order phase transition
like Test Case 1. Therefore, it is tractable to run standard Metropolis-Hastings
MCMC targeting the posterior distribution. In fact, this can be more efficient
because the degree of compression from the prior distribution to the posterior
distribution is very large ($\sim e^{3000}$) in this case.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/summaries.eps}
\end{center}
\caption{Summary images for test case 2. The upper left panel shows the posterior
mean high-resolution scene. The upper right panel shows the posterior mean scene
when observed at the resolution of the data, and the bottom panels show the
model residuals.
\label{fig:summaries}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_N.eps}
\end{center}
\caption{The number of stars in the catalogs produced by \sex~compared with the
actual number of stars in the images. \sex~finds only a small fraction of the
stars actually present. This is to be expected in the high-$\sigma$ case as
\sex~then only searches for stars above a certain flux, but at the low-$\sigma$
end the number of stars is still drastically underestimated. Note that the true
numbers
of stars shown in these plots are $\sim$82\% of the values listed in
Figure~\ref{tab:truth} because here we only count those stars with positions
in $[-1, 1] \times [-1, 1]$.
\label{fig:sex_N}}
\end{figure}
